---
title: "607 HW5 Tidy and Transforment"
author: "Chunmei Zhu"
date: "September 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Data

###chart above describes arrival delays for two airlines across five destinations. Your task is to:
###(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You're encouraged to use a "wide" structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

```{r Import}
# load require packages
require(tidyr)
require(dplyr)
library(plyr)

# Import data from a.csv file
source <-tbl_df(read.csv('D:/CUNY_SPS_DA/607_Data_Aq/week5/a.csv', stringsAsFactors = FALSE)) 
source

# Check data typesraw()
str(source)

```

### Data Cleaning 

```{r a.csv}
# only select rows with complete data in all columns and show data
# renames first two columns 
a <- rename(na.omit(source),c("ï.." ="airlines", "X"="statues"))

# insert missing values in first columns
a[2,1] <- "ALASKA"
a[4,1] <- "AM WEST"

# to shorten the string values to character in second column
#a$statues <- strtrim(a$statues,1)
a
```

###(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

```{r long-formatted-data}
# long-formatted-data
a2 <- a %>%  
  gather(cities, n, Los.Angeles:Seattle, na.rm = TRUE)
a2
```

```{r wide-formatted-data}
# wide-formatted-data
a3 <- a2 %>%
 spread(statues,n)

# correct the column name "on time" to "on_time"
colnames(a3)[4] <- "on_time"

a3
```

```{r AM_WEST-data}
# load requried packages for string value
library(tidyverse)
library(stringr)

# filter data for airline "AM WEST"
am <- a2 %>%
 filter(str_detect(airlines, 'AM WEST'))%>%
 spread(statues,n)
am
```

```{r transfor-row-column-AM_WEST}
# correct the column name "on time" to "on_time"
colnames(am)[4] <- "on_time"

# unite columns airlines and cities
am2<- am %>% 
  unite(AM.city, airlines, cities,sep="." ) %>%
  select(AM.city,delayed,on_time) 

# tranfrom rows to columns
am2<- as.data.frame(t(am2))

# first row as the header in the table a4
colnames(am2) <- unlist(am2[1,])
am2=am2[-1,]
am2
```

```{r  summary-statistics-delayed-AM_WEST}
sub.a3 <- a3%>%  select(airlines,delayed)

sub.a3 %>% summarise(Min = min(delayed, na.rm=TRUE),
                     Median = median(delayed, na.rm=TRUE),
                     Mean = mean(delayed, na.rm=TRUE),
                     Var = var(delayed, na.rm=TRUE),
                     SD = sd(delayed, na.rm=TRUE),
                     Max = max(delayed, na.rm=TRUE))
```

###(3) Perform analysis to compare the arrival delays for the two airlines.
###Answer: AM_WEST airline has more delayed flights than ALASKA airline in every city. Both airlines have the most delayed flights in San.Francisco. 

```{r airlines-delays-analysis}
a3  
df <- a3 %>% 
  mutate(total = on_time + delayed)%>%
  mutate(delayed = delayed/total)%>%
  select(airlines,cities,delayed)%>%
  arrange(cities,airlines)
df

al_delay <- df%>% 
  filter(str_detect(airlines, 'ALASKA'))%>%
  select(cities,delayed)%>%
  arrange(cities)
#al_delay
colnames(al_delay)[2] <-"ALASKA.delayed"

am_delay <- df%>% 
  filter(str_detect(airlines, 'AM WEST'))%>%
  select(cities,delayed)%>%
  arrange(cities)
#am_delay
colnames(am_delay)[2] <-"AM_WEST.delayed"

delayed<- al_delay %>% 
  join(am_delay, by="cities") #%>% 
#  mutate(diff = AM_WEST.delayed - ALASKA.delayed )
delayed


library ('ggplot2')
barplot(delayed$ALASKA.delayed, names.arg=delayed$cities, ylim=c(0,0.30), main="ALASKA Airline Delayed (%)", ylab="blah", xlab="cities", col="darkmagenta")

barplot(delayed$AM_WEST.delayed, names.arg=delayed$cities, ylim=c(0,0.30), main="AM WEST Airline Delayed (%)", ylab="blah", xlab="cities", col="orange")

 matplot(delayed, type = c("b"),pch=1,col = 2:3) #plot
 legend("topleft", legend =1:2, col=1:2, pch=1) # optional legend
 
 

write.table(delayed, "D:/CUNY_SPS_DA/607_Data_Aq/week5/delayed.csv", sep="\t")
 
```

### (4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions. Please include in your homework submission:
###??? The URL to the .Rmd file in your GitHub repository. and
###??? The URL for your rpubs.com web page.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
